Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-14 paper.tex
Minor fixes, added a few todos, and latex comments for conclusion and remaining sections

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    Big Data processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype, as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality, with a number of concurrently processed subject in the  200 ballpark. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Scientific applications differ from typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by a sequence of 
command-line/containerized tools
using a mix of programming languages (C, Fortran, Python, shell 
scripts), referred to as workflows or pipelines. Infrastructure-wise, 
Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

@@

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as 
\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
 We assume that HPC systems are 
multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available. We intentionally did not consider distributed, 
HDFS-like file system, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
become mainstream yet.

@@

The pipelines, benchmarks, performance data, and analysis scripts used 
to implement the methods described hereafter are all available at 
\url{https://github.com/big-data-lab-team/paper-in-mem-locality} for 
further inspection and reproducibility. Links to the pipeline engines 
and processed data are provided in the text.

@@

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. Lazy evaluation is 
represented in Spark through the use of transformations and actions. A series of
transformations may be defined without the data ever being materialized. Using 
this strategy, Spark can make optimizations to ensure optimal processing speeds 
throughout the application.

@@

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping
}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.
%mention results in thunder paper

@@

As Spark is frequently used by the scientific community, we designed 
our experiments using the PySpark API. This is at a cost to performance 
as the PySpark code must undergo Python to Java serialization. We used 
version 2.3.2 installed from \url{https://spark.apache.org}.

@@

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared local filesystem, 
but it may come at the expense of fault-tolerance, as a results of being stored in a disk-local fs. Moreover, the user will need to
ensure that the files are appropriately directed to the nodes that require them as
there is no guarantee of data locality in Nipype.

@@

Page caching is essentially a way to 
emulate in-memory computing at the kernel level, without requiring a 
dedicated engine. The size of the page cache, however, becomes a 
limitation when processes write faster than the disk bandwidth. When 
this happens, the page cache rapidly fills up and writes are limited by 
the disk write bandwidth -- about 500~MB/s for random writes to 
contemporary SSDs  -- as if no page cache was involved.

@@

As in the local disk model, we note that page caching will only be asymptotically
useful when the flush bandwidth is greater than the write throughput of 
the application, that is:
\begin{equation}
\frac{D}{C} \leq \frac{\Delta}{\Gamma}, \label{eq:page-cache-sharedfs}
\end{equation}
where $\Gamma$ is the max number of concurrent processes in the cluster. 
Note that $\frac{\Delta}{\Gamma}$ will usually be much lower than 
$\frac{\delta}{\gamma}$.

@@

 All experiments were executed on  
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell EMC's Zenith 
 cluster}, a Top-500 machine in the Dell EMC HPC Innovation Lab, running 
 Slurm. For the Spark experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
 A 698TB Lustre server was accessible on each node through a 100Gb/s 
 Infiniband network. The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

\subsection{Applications} % 1 page
\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}[1]
    \Input
    \Desc{$x$}{a sleep delay in seconds}
    \Desc{$n$}{a number of iterations}
    \Desc{$C$}{a set of image chunks}
    \Desc{$fs$}{filesystem to write to (mem, tmpfs, local disk, Lustre).}
    \EndInput
    \ForEach{$chunk \in C$}
    \State read $chunk$ from Lustre
    \For{$i \in [1, n]$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
        \If{$i < n$}
        \State save $chunk$ to $fs$
        \EndIf
    \EndFor
    \State write $chunk$ to Lustre
    \EndFor
\end{algorithmic}
\end{algorithm}

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would be read from / written to  Lustre and the final. 
We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality. That is, the chunk were split into partitions, and for each partition of 
chunks, we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, greater chunk size signifies 
a decrease in parallelization, however, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. Unfortunately, 30 chunks could 
no be processed using in-memory Spark as its individual chunk size of 
3GB exceeded the 2GB partition size limit imposed by Spark. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the page cache flushed.
This is explained by the fact that the application data was not large enough to saturate the page  
cache fully (transient phase). As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

Increasing task duration ensured that all file systems had a comparable performance
(Fig.~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
than Spark in-memory at a task duration of 320 seconds, whereas it is 
approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
pattern corroborated our page-cache model which postulates that 
data movement costs will have little impact on compute-intensive tasks. The 
reasoning behind this is that longer tasks give the page cache more time to flush 
between write to disk. The longer the delay, the less likely the page cache is 
to fill up when the concurrent data size written fits within the page cache.

@@

% Comment on the general trend
As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the number 
of chunks increases. This is due to the fact that parallelism increases with a increase in chunk. With 
30 chunks, only 2 CPUs per node are actively working. At 125 chunks, this 
changes to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
active. It can be noted that no significant speedup can be seen between 125 
chunks and 750 chunks. This is because out of the 50 chunks processed by each 
node, only 40 can be processed concurrently, thus requiring two sequential 
incrementation batches and resulting in similar makespans as 125 
chunks.

@@

% Comment on the general trend
Increasing overall data size decreases performance, as can be seen in 
Fig.~\ref{fig:datasize}. When the data size is very small, as is the case 
with the MRI image, all file system makespans are comparable. This is due to the 
fact that page cache can be leveraged fully regardless of file system. However, 
this time, Spark in-memory performed significantly worse than all other 
filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale). Upon further inspection, it appeared that Spark in-memory executed
in a sequential order, on a single worker node. Lack of parallelism for the MRI 
image may be a result of Spark's max partition size, which is by default 128~MB
-- significantly larger than the 13~MB MRI image.

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:modeleval-local}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:modeleval-lustre}
\end{subfigure}
    \caption{Model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

@@

The results also allow us to speculate on the effect of in-memory 
computing on the pre-processing of functional MRI, another typical use 
case in neuroimaging. Assuming an average processing time of 20 minutes 
per subject, which is a ballpark value commonly observed with the 
popular SPM or FSL packages \todo{add refs}, an input data size of 100MB per subject, 
and an output data size of 2GB (20-fold increase compared to input 
size), the D/C ratio would be 1.8MB/s, which would reach the 
$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
that is, if 108 subjects were processed on the same node. This is very 
unlikely as the number of CPUs per node is 40. We therefore conclude 
that in-memory computing is likely to be useless for fMRI analysis.

@@

The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
was found to generally be slower than Spark in all experiments. While a Nipype DAG is 
generated prior to workflow execution, there are no such optimizations, as in 
Spark, to ensure that the least amount of work is performed to produce the 
results.

During the processing of Experiment 3, it was found that Spark split up the data
into an processed it in two batches for the 750 chunk input. That is to say, 
each node in the cluster processed 50 chunks, and since there was only a maximum
of 40 cores available on each node, Spark opted to run to batches of 25 to ensure
a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
Spark opted to perform all the iterations on the first batch (load, increment, save),
and then proceeds to process the second batch. Such an optimization is important, 
even when processing data on disk, as it would presumably increase the occurrance
of cache hits. This may partially explain the speedup seen at 750 chunks in
Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits

@@

Although tmpfs and page cache do improve performance in general, as seen in
Figure~\ref{fig:results}, they do not always perform equivalently to in-memory. 
Tmpfs's main limitation is that data residing on it may be swapped to disk if 
the system's memory usage is high. When it reaches this point, its performance 
slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. As detailed
in~\href{https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/performance_tuning_guide/s-memory-tunables}{RedHatDoc}\todo{add to references},
I/O blocking write to disk occurs when a giving percentage (e.g. 20~\%) of total memory is occupied by dirty data, once this 
threshold is exceeded, processes wanting to execute writes must wait until the dirty data
in the cache is flushed to disk. Furthermore, like memory, tmpfs and page cache are shared resources on a node.
If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
that fill up swap space, this will limit the performance on tmpfs/page cache of other users. 
However, with memory, it is possible to request, through the HPC scheduler, a required
amount of available memory. Ultimately, in-memory data will also need to be spilled to
disk if memory usage exceeds amount of available memory, although disk writes are 
likely to occur in tmpfs and page cache before requested available memory is filled.

@@

A common recommendation in Spark is to limit the number of cores per 
executors to 5, to preserve a good I/O 
throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}. 
We believe that throughput degradation observed with more than 5 cores 
per executor might be coming from full page caches.

@@

@@

\section{Conclusion} % 1 page with refs
\label{sec:conclusion}
% When is it useful to use a Big Data engine?

% "```Disk and network I/O, of course, play a part in Spark performance 
% as well, but neither Spark nor YARN currently do anything to actively 
% manage them.```"
% http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2/

% Future work:
% - scheduling in a shared environment.
% - workflow-aware cache eviction strategies instead of LRU/n

>>>>

\begin{abstract}
    Neuroimaging datasets are rapidly growing in size as a result of 
    advancements in image acquisition methods, open-science and data sharing. 
    However, it remains that current neuroimaging data processing engines do 
    not employ strategies adopted by Big Data processing engines to improve 
    Big Data processing time. Here, we 
    evaluate three Big Data processing stategies (in-memory computing, 
    data-locality and lazy-evaluation) on typical neuroimaging use 
    cases, represented by the BigBrain dataset. We contrast the various 
    processing strategies by using Apache Spark and Nipype, as our 
    representative Big Data and neuroimaging processing engines, on Dell EMC's 
    Top-500 cluster. 
    Big Data thresholds were modeled by comparing the data/compute ratio of the 
    application to the filesystem/concurrent processes ratio. 
    This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, whereas 
when combined with data locality, this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone will
not speed-up current 
functional MRI analyses unless coupled with data 
locality, with a number of concurrently processed subject in the  280 ballpark. 
    In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

Scientific applications differ from typical Big Data use 
cases, which might explain the remaining gap between Big Data and 
scientific engines. While Big Data applications mostly target text 
processing (e.g. Web search, frequent pattern mining, recommender 
systems~\cite{leskovec2014mining}) implemented in consistent software 
libraries, scientific applications often involve 
binary data such as images and signals, processed by a sequence of 
command-line/containerized tools
using a mix of programming languages (C, Fortran, Python, shell 
scripts), referred to as workflows or pipelines. Infrastructure-wise, 
Big Data applications commonly run on 
clouds or dedicated commodity clusters with locality-aware file systems 
such as the Hadoop Distributed File System 
(HDFS~\cite{shvachko2010hadoop}), whereas scientific applications are 
usually deployed on large, shared clusters where data is moved between
data nodes and compute nodes using shared file systems such 
as Lustre~\cite{schwan2003lustre}. Such differences in applications and 
infrastructures have important consequences. To 
mention only one, in-memory computing requires instrumentation to be 
applied to command-line tools.

@@

Infrastructure-wise, we focus on the case of High-Performance Computing 
(HPC) clusters that are typically available through University 
facilities or national computing infrastructures such as 
\href{xsede.org}{XSEDE}, \href{http://computecanada.ca}{Compute Canada} 
or \href{http://www.prace-ri.eu}{PRACE}, as neuroscientists typically use such platforms.
 We assume that HPC systems are 
multi-tenant, that compute nodes are accessible through a batch 
scheduler, and that a file system shared among the compute nodes is 
available. We intentionally did not consider distributed, 
HDFS-like file systems, as initiatives to deploy them in HPC 
centers, for instance Hadoop on-demand~\cite{krishnan2011myhadoop}, have not 
become mainstream yet.

@@

The application pipelines, benchmarks, performance data, and analysis scripts used 
to implement the methods described hereafter are all available at 
\url{https://github.com/big-data-lab-team/paper-in-mem-locality} for 
further inspection and reproducibility. Links to the processing engines 
and processed data are provided in the text.

@@

Two types of operations can be performed on RDDs:
transformations and actions. Applying a transformation to an RDD produces a new
child RDD through a narrow or wide dependency. A narrow dependency signifies 
that the child is only dependent on a single parent partition, whereas a child 
RDD is dependent on all parent partitions in a wide dependency. Examples of 
Spark transformations include map, filter and join. To materialize an RDD, an
action must be performed, such as a reduce or a collect. Lazy evaluation is 
represented in Spark through the use of transformations and actions. A series of
transformations may be defined without the data ever being materialized. Using 
this strategy, Spark can optimize data processing
throughout the application.

@@

Spark will spill RDD data to disk whenever the complete dataset does not fit in
memory. Moreover, as Spark transformations generate new RDDs and numerous 
transformations may occur in a single application, Spark implements a Least-Recently Used (LRU) eviction
policy. If an evicted RDD needs to be reused, Spark will recompute it using the
lineage data collected. As demonstrated in~\cite{freeman2014mapping}, caching significantly improves processing times of iterative algorithms where 
RDDs are reused. It can be of even greater importance if the RDD is costly to 
recompute.
%mention results in thunder paper

@@

As Spark is frequently used by the scientific community, we designed 
our experiments using the PySpark API. This is at a cost to performance 
as the PySpark code must undergo Python to Java serialization. We used Spark
version 2.3.2 installed from \url{https://spark.apache.org}.

@@

In order for Nipype to operate as intended, a filesystem shared by all the nodes 
is required. However, it is still possible to save to a non-shared 
local filesystem, but it may come at the expense of fault-tolerance, as 
a result of being stored in a disk-local file system \todo{I don't get 'as a result of being stored in a disk-local file system'}. Moreover, the 
user will need to ensure that the files are appropriately directed to 
the nodes that require them as there is no guarantee of data locality 
in Nipype.

@@

Page caching is essentially a way to 
emulate in-memory computing at the kernel level, without requiring a 
dedicated engine. The size of the page cache, however, becomes a 
limitation when processes write faster than the disk bandwidth. When 
this happens, the page cache rapidly fills up and writes are limited by 
the disk write bandwidth as if no page cache was involved.

@@

As in the local disk model, we note that page caching will only be 
useful when the flush bandwidth is greater than the write throughput of 
the application, that is:
\begin{equation}
\frac{D}{C} \leq \frac{\Delta}{\Gamma}, \label{eq:page-cache-sharedfs}
\end{equation}
where $\Gamma$ is the max number of concurrent processes \emph{in the cluster}. 
Note that $\frac{\Delta}{\Gamma}$ will usually be much lower than 
$\frac{\delta}{\gamma}$.

@@

 All experiments were executed on 
 \href{https://www.dellemc.com/resources/en-us/asset/sales-documents/products/storage/h16221-hpc-lab-brochure.pdf}{Dell 
 EMC's Zenith cluster}, a Top-500 machine in the 
 \href{https://www.dellemc.com/en-us/solutions/high-performance-computing/HPC-AI-Innovation-Lab.htm}{Dell 
 EMC HPC and AI Innovation Lab}, running Slurm. For the Spark 
 experiments, a Spark cluster was started on our 
 Slurm allocation through access to 16 dedicated nodes. Each compute 
 node ran on RHEL 7.4, was 
 equipped with 40 Intel Xeon Gold 6148 2.4GHz CPUs and had
 187GB of memory, 94GB of tmpfs and a Dell 120GB M.2 SATA SSD as local disk. 
 A 698TB Lustre server was accessible on each node through a 100Gb/s 
 Infiniband network. The apparent write bandwidth of the local disk, 
 Lustre file system and tmpfs were measured by sequentially writing various numbers
 of image blocks containing random intensities, to avoid 
 caching effects (measure\_bandwidth.py script in folder 
 benchmark\_scripts). They are reported in Table~\ref{table:bdwdths}.

@@

\subsection{Applications} % 1 page
\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}[1]
    \Input
    \Desc{$x$}{a sleep delay in seconds}
    \Desc{$n$}{a number of iterations}
    \Desc{$C$}{a set of image chunks}
    \Desc{$fs$}{filesystem to write to (mem, tmpfs, local disk, Lustre).}
    \EndInput
    \ForEach{$chunk \in C$}
    \State read $chunk$ from Lustre
    \For{$i \in [1, n]$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
        \If{$i < n$}
        \State save $chunk$ to $fs$
        \EndIf
    \EndFor
    \State save $chunk$ to Lustre
    \EndFor
\end{algorithmic}
\end{algorithm}

% Incrementation (avoid cache effects in binarization)
In order to investigate how the different strategies impact processing, we 
selected a simple incrementation pipeline (Algorithm~\ref{alg:incrementation}) 
that consisted exclusively of map 
stages. A series of map-only stages would enable us to evaluate the effects of
in-memory computing when data locality is preserved. Incrementation was 
selected 
over other applications, such as binarization, as it ensured that a new image
was created at each step (i.e. no caching effects within the executing 
application). Each partitioned chunk was incremented by 1, in parallel, by
the applications. As incrementing the images is a relatively quick 
process, we added a sleep delay to the incrementation stage in order to 
increase task duration as required. The incremented chunks would be either 
maintained in-memory (Spark only) or saved to either tmpfs, local disk or 
Lustre (Spark and Nipype). Should more than a single iteration be requested, 
the incremented chunks would be incremented again and saved to the same storage. 
This would repeat until the number of requested iterations had elapsed. In all 
conditions, the first input chunks and final output chunks would be read from / written to  Lustre. 
We chose to perform our initial 
reads and final writes on Lustre as a multi-tenant clusters' compute node's 
local storage, if 
available, is generally temporary and only available for the duration of the 
allocation.

@@

We conducted four experiments in which we varied (1) the number of 
iterations in the application ($n$ in 
Algorithm~\ref{alg:incrementation}), (2) the task duration ($x$), (3) 
the chunk size for a constant image size, (4) the total image size. To 
evaluate the page-cache model, experiment conditions fell in different 
regions of Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs}, as summarized in 
Table~\ref{table:experiments}. Among the 16 nodes available, 1 was 
dedicated to the Spark master and driver, and the remaining 15 were 
used as compute nodes. Since data-locality is not normally preserved in 
Nipype (a new Slurm allocation is requested for each processed task), 
we instrumented Nipype to ensure data locality \todo{You could refer to the script doing that in the paper repo.}. That is, the chunk were split into partitions, and for each partition of 
chunks, we requested a Slurm allocation to process the entire 
pipeline in parallel, using Nipype's MultiProc scheduler, on a given 
node. This was possible as no communication was required between the 
processed chunks.

@@

As a third incrementation experiment, we were interested in the effects of 
chunk size on Big Data strategies. Naturally, greater chunk size signifies 
a decrease in parallelization, however, it also signifies an increase in 
sequential I/O (increased $\Delta/\Gamma$ and $\delta/\gamma$). For 
this experiment we partitioned the complete BigBrain image into 30, 125 
and 750 chunks, corresponding to $\gamma$ values of 2, 9 and 25 
respectively. While Spark attempted to load-balance the data, it used 
up only 25 of the 40 cores for 750 chunks. In contrast, Nipype tried to 
use up as many cores as possible. Unlike the previous 
experiment, the D/C ratio was kept static at 178.6MB/s, however, this 
ratio ensured that different regions of the inequality were reached 
depending on amount of parallelization. Unfortunately, 30 chunks could 
no be processed using in-memory Spark as its individual chunk size of 
3GB exceeded the 2GB partition size limit imposed by Spark \todo{You should move this comment to the results or the reader will wonder
what happened to the yellow bar when looking at figure 1.c.}. The number of iterations was
fixed to 10, and the task duration was adjusted so that C remained constant at 4,400s.

@@

Fig.~\ref{fig:iterations} shows the difference between the different 
filesystem choices given the number of iterations. At 1 iteration, all 
filesystems behave the same, although the application was writing faster than the page cache flushed. \todo{Update the graph to the same y scale and 'broken'
bars, and check that this impression still holds...}
This is explained by the fact that the application data was not large enough to saturate the page  
cache fully (transient phase). As the total amount of data written by the 
application increases to 750~GB, there is a greater disparity between 
Lustre and in-memory (2.67~x slower, on average). Local disk 
performance, however, 
is still comparable to memory (1.38~x slower, on average). Despite local disk and 
Lustre both being in transient state, local disk encounters less contention 
than what would be found on Lustre.

@@

Increasing task duration ensured that all file systems had a comparable performance
(Fig.~\ref{fig:cputime}). Lustre, for instance, is approximately 1.01x slower
than Spark in-memory at a task duration of 320 seconds, whereas it is 
approximately 3.25x slower that Spark in-memory with 2.4 second tasks. This 
pattern corroborated our page-cache model which postulates that 
data movement costs will have little impact on compute-intensive tasks. The 
reasoning behind this is that longer tasks give the page cache more time to flush 
between disk writes. The longer the delay, the less likely the page cache is 
to fill up when the concurrent data size written fits within the page cache.

@@

% Comment on the general trend
As can be seen in Fig.~\ref{fig:numchunks}, makespan decreases when the 
number of chunks increases. This is due to the fact that parallelism 
increases with an increase in the number of chunks. With 30 chunks, 
only 2 CPUs per node are actively working. At 125 chunks, this changes 
to a maximum of 9 CPUs per node, and at 750 chunks all 40 CPUs can be 
active. It can be noted that no significant speedup can be seen between 125 
chunks and 750 chunks \todo{This is not true for in-memory, Spark tmpfs and Spark local disk}. This is because out of the 50 chunks processed by each 
node, only 40 can be processed concurrently, thus requiring two sequential 
incrementation batches and resulting in similar makespans as 125 
chunks.

@@

% Comment on the general trend
Increasing overall data size decreases performance, as can be seen in 
Fig.~\ref{fig:datasize}. When the data size is very small, as is the case 
with the MRI image, all file system makespans are comparable. This is due to the 
fact that page cache can be leveraged fully regardless of file system. However, 
this time, Spark in-memory performed significantly worse than all other 
filesystems, with a makespan of 2,211 seconds (not reported on the graph to preserve the y scale) \todo{You may want to add it back when you have
broken bars}. Upon further inspection, it appeared that Spark in-memory executed
in a sequential order, on a single worker node. Lack of parallelism for the MRI 
image may be a result of Spark's max partition size, which is by default 128~MB
-- significantly larger than the 13~MB MRI image.

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:modeleval-local}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:modeleval-lustre}
\end{subfigure}
    \caption{Page cache model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

@@

The results also allow us to speculate on the effect of in-memory 
computing on the pre-processing of functional MRI, another typical use 
case in neuroimaging. Assuming an average processing time of 20 minutes 
per subject, which is a ballpark value commonly observed with the 
popular SPM or FSL packages \todo{add refs}, an input data size of 100MB per subject, 
and an output data size of 2GB (20-fold increase compared to input 
size), the D/C ratio would be 1.8MB/s, which would reach the 
$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
that is, if 108 subjects were processed on the same node. This is very 
unlikely as the number of CPUs per node is 40. We therefore conclude 
that in-memory computing is likely to be useless for fMRI analysis.
Naturally, this estimate is strongly dependent on the characteristics of the cluster.

@@

The effects of lazy-evaluation can be seen throughout the experiments. Nipype 
was found to generally be slower than Spark in all experiments. While a Nipype DAG is 
generated prior to workflow execution, there are no such optimizations, as in 
Spark, to ensure that the least amount of data transfer is performed to produce the 
results.

During the processing of Experiment 3, it was found that Spark split up the data
into an processed it in two batches for the 750 chunk input. That is to say, 
each node in the cluster processed 50 chunks, and since there was only a maximum
of 40 cores available on each node, Spark opted to run \todo{a word is missing here} to batches of 25 to ensure
a balanced workload. Rather than running each iteration on the full dataset, as with Nipype, 
Spark opted to perform all the iterations on the first batch (load, increment, save),
and then proceeds to process the second batch. Such an optimization is important, 
even when processing data on disk, as it would presumably increase the occurrence
of cache hits. This may partially explain the speedup seen at 750 chunks in
Figure~\ref{fig:numchunks}.% Spark local vs nipype local: presumably increases cache hits
\todo{Add as sentence on this paragraph in the abstract.}

@@

Although tmpfs and page cache do improve performance in general, as seen in
Figure~\ref{fig:results}, they do not always perform equivalently to in-memory. 
Tmpfs's main limitation is that data residing on it may be swapped to disk if 
the system's memory usage is high. When it reaches this point, its performance 
slows down to swap disk bandwidth, as observed in Figure~\ref{fig:datasize}. Page cache suffers a similar dilemma. As detailed
in~\href{https://access.redhat.com/documentation/en-us/red_hat_enterprise_linux/6/html/performance_tuning_guide/s-memory-tunables}{RedHatDoc}\todo{add to references},
I/O blocking write to disk occurs when a giving percentage (e.g. 20~\%) of total memory is occupied by dirty data, once this 
threshold is exceeded, processes wanting to execute writes must wait until the dirty data
in the cache is flushed to disk \todo{Revise this sentence (grammar?)}.

Furthermore, like memory, tmpfs and page cache are shared resources on a node.
If users on the node are heavily using memory to incite tmpfs to write to swap space, or are performing data-intensive operations
that fill up the page cache, this will limit the performance on tmpfs/page cache of other users. 
However, with memory, it is possible to request, through the HPC scheduler, a required
amount of available memory. Ultimately, in-memory data will also need to be spilled to
disk if memory usage exceeds amount of available memory, although disk writes are 
likely to occur in tmpfs and page cache before requested available memory is filled.

@@

A common recommendation in Spark is to limit the number of cores per 
executors to 5, to preserve a good I/O 
throughput\footnote{\url{http://blog.cloudera.com/blog/2015/03/how-to-tune-your-apache-spark-jobs-part-2}}, but the rationale
for this recommendation is hardly explained. 
We believe that throughput degradation observed with more than 5 cores 
per executor might be coming from full page caches.

@@

% We haven't benchmarked shuffles but we know it's important.

@@

\section{Conclusion} % 1 page with refs
\label{sec:conclusion}

% Summary of the achievements.
% When is it useful to use a Big Data engine for neuroinformatics applications?
